Sampling requirements to detect ecosystem change

### 2020-04-07 R/supp_material.Rmd
change title
correct title
Changed order of figures, added figure, and edited text

\begin{center}
\textbf{\Large Supplementary Material for: Sampling requirements and approaches to detect ecosystem shifts}
\vspace{7 mm}

\textsc{Rosalie Bruel$^{1}$ and Easton R. White$^{2}$}
\vspace{5 mm}

\normalsize{\indent $^{1}$Rubenstein Ecosystem Science Laboratory, University of Vermont, VT 05405, USA \\ $^{2}$Department of Biology, University of Vermont, VT 05405, USA}
\end{center}

@@

# Parameter sensitivity

We investigated how different parameters change the power of the analysis, and found that autocorrelation at lag-1 (phi) had little impact on the power of the analysis amplitude of the shift was the most important determinant of the power of the analysis. For null standard deviation, the analysis performed well independently of the shift size or autocorrelation. When standard deviation increased, the power of the analysis was higher for null autocorrelation (Fig. \ref{fig:parameter_sensitivity_plot}).

```{r parameter_sensitivity,message=F,warning=F,echo=F,eval=F,cache=F}
# Build 1x3 plot with sigma vs phi with color equal to distance to changepoint, power, or min time. The three panel plots would represent different shift size values

source('R/sampling_regular.R')
source('R/sampling_random.R')
source('R/build_time_series.R')
source('R/calc_min_time2.R')
require(tidyverse)

number_sims <- 10
# Find minimum time for different parameter combos

sigma_vec = seq(0.01,0.3,0.02)
phi_vec = seq(-0.8,0.8,0.1)
shift_size_vec = c(0.1,0.4,0.8)

output_theory_approach <- data.frame(expand.grid(sigma_vec,phi_vec,shift_size_vec))
names(output_theory_approach) <- c('sigma_vec', 'phi_vec', 'shift_size_vec')
output_theory_approach$power = 0

#params <- tibble(shift_time = sample(30:70,max(simulation_number),replace = T))

shift_time = sample(30:70,number_sims,replace=T) # Keep shift times constant for each param combo

# Run through params
for (i in 1:nrow(output_theory_approach)){

# Build XX number of time series
params <- list(sigma=rep(output_theory_approach$sigma_vec[i],number_sims),phi=rep(output_theory_approach$phi_vec[i],number_sims),shift_size=rep(output_theory_approach$shift_size_vec[i],number_sims),shift_time = shift_time)

time_series <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) sample_regular(df,20)$changepoint)

power <- sum(abs(time_series-shift_time)<5)/number_sims

#print(i)
output_theory_approach$power[i] <- power
}

write.csv(x = output_theory_approach,file = 'Output/parameter_sensitivity.csv',quote = FALSE )

require(ggplot2)
require(viridis)
ggplot(aes(x=phi_vec,y=sigma_vec,fill=power),data=output_theory_approach) + geom_tile() +
   scale_fill_viridis(discrete=FALSE,begin = 0.1,end=0.9) +
  facet_wrap(~shift_size_vec) +
   labs(fill='Statistical power') +
   xlab("Lag-1 autocorrelation (phi)") +
   ylab("Standard deviation (sigma)") +
   theme_classic(base_size = 14, base_family = "")

```{r parameter_sensitivity_plot,echo=F,message=F,warnings=F,fig.height=4,fig.cap='Regular sampling statistical power (fraction of 100 simulations which detecting a changepoint within five time points of the true changepoint) for different levels of standard deviation, lag-1 autocorrelation, and shift size.\\label{fig:parameter_sensitivity_plot}'}
output_theory_approach <- read.csv(file = 'Output/parameter_sensitivity.csv')

require(ggplot2)
require(viridis)
ggplot(aes(x=phi_vec,y=sigma_vec,fill=power),data=output_theory_approach) + geom_tile() +
   scale_fill_viridis(discrete=FALSE,begin = 0.1,end=0.9) +
  facet_wrap(~shift_size_vec) + 
   theme_classic(base_size = 14, base_family = "")+
  theme(plot.subtitle = element_text(hjust = 0.5)) +
   labs(fill='Power',subtitle = 'Shift size') +
   xlab("Lag-1 autocorrelation (phi)") +
   ylab("Standard deviation (sigma)")
```

@@

source('R/sampling_iterative.R')
source('R/build_time_series.R')
source('R/updated_param_sensitivity.R')
source('R/calc_min_time2.R')
require(tidyverse)

@@

for (param_num in 11:nrow(sigma_output)){
  sigma_output$min_time[param_num] <- calc_min_time('sigma',sigma_output$sigma_vec[param_num],'iterative')
  print(sigma_output)
}

>>>>

\begin{center}
\textbf{\Large Supplementary material for: Sampling requirements and approaches to detect ecosystem shifts}
\vspace{5 mm}

\textsc{Rosalie Bruel$^{*}$\footnote{*The authors contributed equally to this work. $\dagger$Corresponding author (eastonrwhite@gmail.com)} and Easton R. White$^{* \dagger 2,3}$}
\vspace{3 mm}

\normalsize{\indent $^{1}$Rubenstein Ecosystem Science Laboratory, University of Vermont, VT, USA \\ $^{2}$Department of Biology, University of Vermont, VT, USA \\
$^{3}$Gund Institute for Environment, University of Vermont, VT, USA}
\end{center}

@@

<!--
# Parameter sensitivity

We investigated how different parameters change the power of the analysis, and found that autocorrelation at lag-1 (phi) had little impact on the power of the analysis amplitude of the shift was the most important determinant of the power of the analysis. For null standard deviation, the analysis performed well independently of the shift size or autocorrelation. When standard deviation increased, the power of the analysis was higher for null autocorrelation (Fig. \ref{fig:parameter_sensitivity_plot}).
-->

@@

source('R/sampling_iterative.R')
source('R/build_time_series.R')
#source('R/updated_param_sensitivity.R')
source('R/calc_min_time2.R')
require(tidyverse)

@@

for (param_num in 1:nrow(sigma_output)){
  sigma_output$min_time[param_num] <- calc_min_time('sigma',sigma_output$sigma_vec[param_num],'iterative')
  print(sigma_output)
}
